### 2019-10-17 tex/figures/traq_illustration_too_much.tex

\input{traq_common_beginning.tex}
%\fill[pattern=flexible hatch, pattern color=red, hatch distance=1mm, hatch thickness=0.01mm] (0,0) rectangle ++(2.5,-0.5);
\draw[draw opacity=0,fill=red,fill opacity=0.3] (0,0) rectangle ++(2.5,-0.5);
%\draw[name path=A] (0,-0.5)-- ++(2.5, 0);

\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (0.75,-0.5) node [midway,left,align=right,font=\fontsize{3pt}{0}\selectfont] {standing\\ queue};
%\draw[densely dotted] (0.75,0) -- (0.75,-0.5);

\draw[dashed] (0,-0.5) -- ++(2.5,0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.95,-0.25) (optimal) {optimal\\ window};
\draw[->, line width=0.05mm] (optimal) -- ++(0,-0.25);

\input{traq_common_end.tex}